The Forward March of Categorical Tolerance in the United States\thanks{Reserved for acknowledgments}

### 2025-11-14 manuscript.tex
Updates from Overleaf

\newpage
\begin{abstract}
\end{abstract}
\newpage
\section*{Introduction}
Ever since Richard Peterson introduced the felicitous term ``patterns of cultural choice'' into sociological discourse more than four decades ago \citep{peterson1983patterns-f85}, sociologists have been busy empirically uncovering and debating the sociological reality, drivers, and correlates of various such patterns. These include, most famously and influentially, the Petersonian cultural \textit{omnivore} \citep{peterson1996changing-967, goldberg2011mapping-77a, warde2008omnivorous-e77} along with their evil twin the culturally exclusive \textit{univore} \citep{bryson1997what-933, peterson1992understanding-ce3, tonkovi2025globallocal-96f}, but also culturally \textit{voracious} consumers \citep{sullivan2007omnivore-018}, \textit{poly-purists} and \textit{poly-mixers} \citep{goldberg2016what-4e3, poecke2018pure-873}, \textit{decisives} and \textit{indecisives} \citep{sonnett2004musical-961}, and even plain old \textit{snobs} engaging in (presumably) Bourdieusian boundary-drawing distinction \citep{atkinson2011context-d44, nault2021social-1bf}, among many others.

In a paper published almost a decade ago, \citet{lizardo2016end-4fb}---hereafter L\&S---introduced another such pattern of cultural choice, which they labeled \textit{categorical tolerance}, to refer to the phenomenon of a growing proportion of the population who refuse to report dislikes of \textit{any} cultural genres---typically musical genres---in standard social science surveys designed to measure patterns of cultural taste. In that paper, L\&S noted that categorical tolerance appeared to be on the rise in the United States, going from about five percent of the population in the canonical 1993 General Social Survey culture module dataset to about sixteen percent in a then-recent survey of a representative sample of Americans collected in 2012, which matched the GSS 1993 battery of items on likes and dislikes of musical tastes across eighten genres. This resulted in a significant shift in the shape of the univariate count distributions of number of dislikes, which went from a standard Poisson distribution in the GSS 1993 to a count distribution with excess zeroes in 2012 \citep[p.90, Figure 1]{lizardo2016end-4fb}.

@@

This paper uses recently collected data from a representative sample of the American population to update the empirical picture of categorical tolerance in the United States for the third decade of the twenty-first century. First, I ask whether we can observe a continued rise in the categorical tolerance effect in the more recent data. Secondly, I examine the socio-demographic correlates of categorical tolerance to determine whether this pattern of cultural choice persists in a similar socio-demographic ``niche'' \citep{mcpherson2004blau-9fe} as L\&S observed in 2012, or whether its location in social space has evolved since then. Third, I also examine the socio-demographic correlates of \textit{symbolic exclusion} among people at risk of expressing dislikes \citep{bryson1996anything-311}, to see if we can observe similar patterns to those previously uncovered by L\&S. Finally, the paper adds a missing dimension to previous work on categorical tolerance and symbolic exclusion by investigating whether these cultural taste patterns are connected to political identification as liberal or conservative, given recent work indicating that cultural taste patterns have themselves become embroiled in the ``oil spill'' produced by the continuing political polarization of the American mass publics in the last two decades. This phenomenon has accelerated in the Trump era \citep{dellaposta2020pluralistic-ed4, rawlings2023polarization-0af}.

To preview the main results: I find that categorical tolerance has indeed continued its onward march among Americans, with about twenty percent of respondents failing to report disliking any of the twelve genres in data collected from two survey samples in two separate platforms. Secondly, the social location of categorical tolerance is fairly similar to that observed by L\&S a decade ago; younger adults are overrepresented among categorical tolerants relative to older adults. Similarly, as L\&S found in 2012, non-white respondents are overrepresented among categorical tolerants in the United States. One distinct difference between the current findings and L\&S's results, however, is a seeming trickle-up of categorical tolerance as an elite taste pattern. Those reporting the highest levels of education are overrepresented in this group, indicating that Bourdieusian status display dynamics may increasingly be at play for this pattern of taste. Finally, while categorical tolerance is not polarized along liberal/conservative lines, \textit{symbolic exclusion} patterns net of categorical tolerance are. Adjusting for other socio-demographic predictors, conservatives are more likely to express a larger number of cultural dislikes than liberals. This indicates that political ideology, along with age---but not education---is now the most reliable predictor of expressing dislikes for cultural genres in the United States today.

The rest of the paper is organized as follows. The next section briefly reviews recent work that builds on L\&S's idea of categorical tolerance, as well as other work looking at the increasing linkages between cultural taste and political ideology. The next sections introduce the data and outline the basic analytic plan, presenting the main results. Finally, the discussion section summarizes the key findings, highlights their main implications, and suggests avenues for future work.

@@

The general empirical implication is that cultural omnivores should be more (explicitly) tolerant, liberal, and inclusive, displaying a fundamental openness to consume and engage in a wide variety of cultural practices and experiences. However, the Petersonian thesis does not imply that omnivores engage in an indiscriminate acceptance of all cultural forms. Instead, subsequent research goes through some pains to emphasize that cultural omnivores still engage in ``patterned tolerance" and selective exclusion of low-status cultural forms \citep{bryson1996anything-311, lindblom2019anything-530}. This means that while cultural elites embrace a wide variety of cultural genres, they often express explicit dislikes toward a few specific genres, particularly those associated with lower-status groups \citep{bryson1996anything-311, lindblom2022growing-ded}.

Why are dislikes expressed in social surveys on cultural tastes important? As \citet{bourdieu1984distinction-835} himself noted, taste is often asserted negatively through refusal (p. 56). \citet{bryson1996anything-311} linked this dynamic to the idea of ``symbolic exclusion'' or the \textit{indirect} action of drawing a symbolic (social) boundary against particular (usually lower status) groups by distancing oneself from the cultural objects that those groups are perceived to consume \citep{lizardo2016cultural-aaa}. This recognition has led scholars to stress the critical importance of studying ``dislikes" alongside ``likes" to fully comprehend the dynamics of symbolic exclusion and the construction of cultural boundaries \citep{lizardo2015musical-8c6, lindblom2022growing-ded, oncini2023cultural-323}. This general framework has proven helpful in understanding social and cultural dynamics even outside the case of the United States. For example, a recent study of musical taste conducted in South Korea found that tolerant omnivores still excluded genres like ``gugak" (Korean traditional music) and ``trot," (one of the oldest forms of Korean popular music) which are preferred by lower-educated and older individuals, suggesting that the more ``cosmopolitan'' omnivores still engage in patterns of symbolic exclusion directed at older persons of lower educational status \citep[321]{kim2020anything-71b}.

@@

Various studies conducted since Bryson's consistently show an empirical link between cultural openness and social tolerance across multiple contexts. For instance, \citet{cho2023are-61e} show that cultural omnivores in South Korea are more accepting of immigrants, demonstrating a greater willingness to include them as friends, neighbors, and even marriage partners. Beyond immigration, cultural omnivorousness exhibits a positive correlation with a liberal political outlook and a higher degree of social trust, particularly towards outgroup members. Their diverse cultural engagements are also associated with a cosmopolitan outlook, support for social integration and inclusiveness, and progressive values such as the rejection of traditional gender roles and acceptance of alternative sexual lifestyles \citep{chan2019understanding-1d0}. In line with this work, A study conducted in Finland by \citep{lindblom2022growing-ded} indicates that omnivores hold very liberal views on immigration, law enforcement, and same-sex relationships, while prioritizing environmental protection \citep{lindblom2022growing-ded}.

Conversely, Univorism, a concept first conceptualized by \citet{peterson1992understanding-ce3} and empirically studied by \citet{bryson1997what-933}, describes individuals characterized by narrow and limited cultural tastes, typically confining themselves to a smaller range of popular cultural practices \citep{lindblom2022growing-ded, cho2023are-61e}. This restricted taste profile is often characteristic of older age groups and those with lower educational attainment \citep{sokolova2020does-65c, lindblom2019anything-530}. Recent work shows that univores exhibit less liberal and less tolerant attitudes across various social and political domains. For instance, univores tend to be positioned on the conservative end of the liberal-conservative spectrum, are more skeptical towards fellow citizens, indicating lower social trust,  have lower voter turnout, and espouse more conservative political views that are critical of immigration and generally do not prioritize environmental protection \citep{lindblom2022growing-ded, cho2023are-61e, chan2019understanding-1d0}.

@@

\section*{Data}
The data for this study come from three rounds of data collection on the cultural tastes of Americans. The two waves of data collection resulted in two convenience samples: one collected during the summer of 2021 and the other during the summer of 2023, with an oversample of people with an advanced educational degree. Both of these samples come from the Prolific Academic (ProA) platform. Prolific has emerged as a preferred alternative to Amazon’s Mechanical Turk for researchers seeking high-quality, suitably motivated participant pools. Unlike MTurk, ProA was explicitly designed with social-scientific researchers in mind and offers a high-quality and comparatively better-compensated participant pool. Also, unlike MTurk, ProA mandates a minimum payment based on the expected time for task completion. Participants in our study were compensated at a rate of \$9.50 per hour, which ProA indicates as a “good” compensation rate. The main participant pool for this study on the Prolific platform was limited to participants between the ages of eighteen and eighty-five who reside in the United States and whose first language is English (or are fluent in English). The study was also restricted to participants who scored at or above a 95\% in ProA’s ``approval rate'' (the proportion of time a participant’s study response has been accepted in other studies as adequate). Survey completion was verified by embedding a completion code as the last question, which participants had to input on the Prolific website to confirm they had read and completed each survey question.

In the summer of 2025, I collected a third sample to try to replicate the results of the first two waves on a more representative sample of American respondents. For this purpose, I used the recently launched Cint (previously Lucid) Theorem platform, which provides access to a quota-sampled set of respondents matched to the U.S. census by age, gender, ethnicity, and region. Like Prolific, Cint Theorem is designed primarily for social science research. Cint employs a ``marketplace'' approach to recruiting study participants by connecting researchers to various suppliers and offering a variety of incentives for respondents. In this way, Cint provides relatively affordable access to participants for researchers (about \$1.50 per completed survey) and high-quality samples \citep{crabtree2025yes-e5f}. Cint also uses various automated screens (e.g., invalid traffic, duplicate IPs) and response quality indicators (e.g., CAPTCHA scores) to weed out bots and prevent less attentive participants from accessing the survey. Representative samples of the U.S. population from Cint Theorem have been used to good effect in recent social scientific work on the political and social attitudes of Americans (e.g., Keskintürk et al. 2025).

@@

\section*{Results}
\subsection*{Probability of Dislikes by Genre}
Table~\ref{tab:dis-by-genres} shows the probability of any genre experiencing a judgment of dislike for both the Prolific and Lucid samples, pooling the 2021 and 2023 Prolific samples. As we can see, Metal is the genre more likely to be disliked in both samples, consistent with previous work on the musical tastes of Americans based on data collected at earlier time points \citep{bryson1996anything-311, lizardo2015musical-8c6}. Both samples also agree that the two genres least likely to be disliked are Rock and Pop music. There is some disagreement across samples on the exact ranking of other genres. Country music experiences the most significant disparity, being the second most disliked genre in the Prolific sample but close to the bottom in the Lucid sample. Beyond that, both datasets agree that Musicals, Rap, and Disco are at the top of the dislike list, while Classical, R\&B, and Jazz are closer to the bottom.

@@

\subsection*{Univariate Distribution of Number of Dislikes}
Table~\ref{tab:dis-dist} shows the univariate distribution of the number of dislikes for both the Prolific and Lucid samples. As we can see, both datasets agree closely in their estimate of the proportion of categorical tolerants, with the Prolific data returning an estimate of about 19\% of the sample and the Lucid dataset returning an estimate of about 20\%. When we combine both, we get an estimate of 19.6\%, suggesting that about a fifth of the American population today can be classified as categorical tolerants, an increase from L\&S's estimate of about 15\% using data from 2012.\footnote{We should be cautious when making direct comparisons, given that the number of genres included is smaller (twelve versus twenty in L\&S's 2012 survey) in both the Prolific and Lucid samples. Nevertheless, we can conclude that categorical tolerance remains a pervasive phenomenon among a substantial number of Americans.}

@@

\begin{figure}[ht!]
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol.png}
    \caption{Univariate plot of number of dislikes, joint Prolific and Lucid samples (N = 2306).}
    \label{fig:main}
\end{figure}

@@

\subsection*{Multivariable Analysis Results}
In their original paper, \citet{lizardo2016end-4fb} proposed modeling categorical tolerance and symbolic exclusion jointly as a hurdle process. The basic idea is that categorical tolerants are a distinct slice of the population who are not at risk of engaging in symbolic exclusion (expressing dislikes). Thus, we can statistically model the joint predictors of both tendencies by assuming that there is a set of predictors that drive the probability of being in the categorical tolerant group (thus not being at risk of expressing dislikes), and net of this, there is a second set of predictors that drive the number of dislikes expressed for those who are at risk of doing so. The zero-inflated Poisson model is thus a natural statistical framework for modeling this dual regime data-generating process \citep{zorn1998analytic-6b1}. Since the univariate distribution of cultural dislikes is comparable across samples (see ~\ref{tab:dis-dist}), I pool the three samples to increase statistical power and report results from the pooled sample analysis, including a binary indicator for each survey sample (using the 2021 Prolific sample as the reference category) in each model. I use listwise deletion of cases with missing values for the socio-demographic covariates included in the model, which leaves us with a total analytical sample of $N = 2091$.

@@

In line with this analytic strategy, Table~\ref{tab:zinf} shows the coefficient estimates from a set of four zero-inflated Poisson regression models. The specifications in each model, for both the count and zero-inflated components, are informed by the descriptive patterns we considered in Figures~\ref{fig:cat-tol} and~\ref{fig:grd-int}. Coefficient estimates corresponding to the count part of the model---predicting the number of dislikes among those who are at risk of expressing them---are shown in the top half of the table. Coefficient estimates for the zero-inflation part of the model, predicting the odds of landing in the categorical tolerant group---and thus not being at risk of engaging in symbolic exclusion---are presented in the lower half of the table. In all models, age, education, income, and political ideology (with higher values indicating a more liberal stance) are entered as ordered numerical inputs. Race (White) and Gender (Female) are entered into the model as dummy indicators. The zero-inflated specifications also contain squared terms for education and income to model expected curvilinear effects (see Figures~\ref{fig:cat-tol-edu} and~\ref{fig:cat-tol-inc}). Finally, I include indicator variables for each of the surveys (one for the 2023 Prolific sample and another for the 2025 Lucid sample), with the Prolific survey conducted in 2021 as the reference category.

Let us begin with the predictors of symbolic exclusion. Model one includes all of the main socio-demographic predictors. The results align with expectations derived from the descriptive patterns considered earlier. Age increases the likelihood of engaging in symbolic exclusion, but education and income reduce this tendency. Nevertheless, the education effect, as suggested by Figure~\ref{fig:cat-tol-edu}, is substantively small and not statistically significant ($p = 0.09$) using the traditional cutoff value of $p < 0.05$, and becomes smaller and crosses the threshold to decidedly not statistically significant in later specifications. Being politically liberal, on the other hand, has a substantively significant effect, one that is robust---in terms of statistical significance---across models. As suggested by Figure~\ref{fig:cat-tol-pol}, self-identified Liberals are less likely to engage in symbolic exclusion than conservatives. These results indicate that in the United States today, political liberalism is a more reliable predictor of whether a person will engage in symbolic exclusion than is educational attainment, a situation quite different from that which obtained in the 1990s \citep{bryson1996anything-311}. When it comes to ethnoracial identity, self-identified white respondents are more likely to express dislikes compared to respondents who identified with one of the non-white ethnoracial categories ($p < 0.01$). Finally, there is no appreciable gender-identity difference in the propensity to engage in symbolic exclusion via cultural dislikes ($p = 0.52$).

Model 1 includes age as the sole predictor in the zero-inflated part of the model, providing strong evidence that categorical tolerance is primarily a cultural habit among the young and is less prevalent among older respondents ($p < 0.001$). Model 2 includes race and gender as additional predictors of categorical tolerance, verifying the expectation that white respondents are less likely to be found in this group (compared to non-white respondents) and women are less likely to be categorical tolerants (compared to respondents who identify as men). Model 3 introduces education as an additional predictor of categorical tolerance, including a linear input for the respondent's education and a square term for the same covariate. The results verify the idea, suggested by Figure~\ref{fig:cat-tol-edu}, the education effect on the odds of being a categorical tolerant is parabolic, with a negative linear component (indicating that lower education person are more likely to be in the categorical tolerant group) and a positive squared component (indicating that at some point the negative effect reverses, and very high-education individuals end up over-represented among the categorical tolerants).

That said, the linear component of education is of only marginal statistical significance in Model 3 ($p = 0.07$). This suggests that the bulk of the systematic effect of education is contained in the positive quadratic component, indicating a larger propensity of the very highly educated to be over-represented among categorical tolerants. This impression is confirmed in Model 4, which adjusts for the non-linear effect of self-reported family income on the odds of being a categorical tolerant. Once we do that, the main negative linear effect of education is attenuated ($p = 0.22$), but the exponential positive effect remains ($p < 0.05$), using statistical significance as a criterion. Both components of the income effect, on the other hand, have a statistically significant impact on the odds of being classified as a CT, with a negative linear effect and a positive quadratic effect, suggesting an under-representation of people in middle-income categories among the categorically tolerant.

@@

\subsection*{Summary of Key Results and Limitations}

>>>>

\newpage
\begin{abstract}
This paper updates the empirical picture of Categorical Tolerance (CT), namely, the pattern of refusing to report dislikes across cultural genres, in the United States during the third decade of the twenty-first century. Analyzing recent survey data from two platforms, I find that CT has continued its march among Americans, reaching approximately 20\% of respondents. The analysis confirms the persistence of earlier-observed demographic trends, showing that CT is strongly associated with younger cohorts and non-white individuals. However, I also find that individuals reporting the highest educational attainment are now overrepresented among categorical tolerants, suggesting that CT may increasingly function as an elite cultural strategy consistent with contemporary forms of status display, signaling openness and refusal to refuse. Furthermore, the study uses political ideology as a predictor of symbolic exclusion, revealing that while the odds of being a CT itself are not strongly polarized, the inclination toward symbolic exclusion among non-CTs is, with conservatives significantly more likely to express a greater volume of cultural dislikes than liberals. This result shows that political ideology, alongside age and race, is a central determinant of cultural exclusion in the contemporary context. 
\end{abstract}
\newpage
\section*{Introduction}
Ever since Richard Peterson introduced the felicitous term ``patterns of cultural choice'' into sociological discourse more than four decades ago \citep{peterson1983patterns-f85}, sociologists have been busy empirically uncovering and debating the sociological reality, drivers, and correlates of various such patterns. These include, most famously and influentially, the Petersonian cultural \textit{omnivore} \citep{peterson1996changing-967, goldberg2011mapping-77a, warde2008omnivorous-e77} along with their evil twin the culturally exclusive \textit{univore} \citep{bryson1997what-933, peterson1992understanding-ce3, tonkovi2025globallocal-96f}, but also culturally \textit{voracious} consumers \citep{sullivan2007omnivore-018}, \textit{poly-purists} and \textit{poly-mixers} \citep{goldberg2016what-4e3, poecke2018pure-873}, \textit{decisives} and \textit{indecisives} \citep{sonnett2004musical-961}, and even plain old \textit{snobs} engaging in (presumably) Bourdieusian boundary-drawing distinction \citep{atkinson2011context-d44, nault2021social-1bf}, among many others.

In a paper published almost a decade ago, \citet{lizardo2016end-4fb}---hereafter L\&S---introduced another such pattern of cultural choice, which they labeled \textit{categorical tolerance}, to refer to the phenomenon of a growing proportion of the population who refuse to report dislikes of \textit{any} cultural genres---typically musical genres---in standard social science surveys designed to measure patterns of cultural taste. In that paper, L\&S noted that categorical tolerance appeared to be on the rise in the United States, going from about five percent of the population in the canonical 1993 General Social Survey culture module dataset to about sixteen percent in a then-recent survey of a representative sample of Americans collected in 2012, which matched the GSS 1993 battery of items on likes and dislikes of musical tastes across eighteen genres. This resulted in a significant shift in the shape of the univariate count distributions of number of dislikes, which went from a standard Poisson distribution in the GSS 1993 to a count distribution with excess zeroes in 2012 \citep[p.90, Figure 1]{lizardo2016end-4fb}.

@@

This paper uses recently collected data from a representative sample of the American population to update the empirical picture of categorical tolerance in the United States for the third decade of the twenty-first century. First, I ask whether we can observe a continued rise in the categorical tolerance effect in the more recent data. Secondly, I examine the socio-demographic correlates of categorical tolerance to determine whether this pattern of cultural choice persists in a similar socio-demographic ``niche'' \citep{mcpherson2004blau-9fe} as L\&S observed in 2012, or whether its location in social space has evolved since then. Third, I also examine the socio-demographic correlates of \textit{symbolic exclusion} among people at risk of expressing dislikes \citep{bryson1996anything-311}, to see if we can observe similar patterns to those previously uncovered by L\&S. Finally, the paper adds a missing dimension to previous work on categorical tolerance and symbolic exclusion by investigating whether these cultural taste patterns are connected to political identification as liberal or conservative, given recent work indicating that cultural taste patterns have themselves become embroiled in the ``oil spill'' produced by the continuing political polarization of the American mass public in the last two decades. This phenomenon has accelerated in the Trump era \citep{dellaposta2020pluralistic-ed4, rawlings2023polarization-0af}.

To preview the main results: I find that categorical tolerance has indeed continued its onward march among Americans, with about 20\% of respondents failing to report disliking any of the 12 genres in data collected from two survey samples on two separate platforms. Secondly, the social location of categorical tolerance is fairly similar to that observed by L\&S a decade ago; younger adults are overrepresented among categorical tolerants relative to older adults. Similarly, as L\&S found in 2012, non-white respondents are overrepresented among categorical tolerants in the United States. One distinct difference between the current findings and L\&S's results, however, is a seeming trickle-up of categorical tolerance as an elite taste pattern. Those reporting the highest levels of education are overrepresented in this group, indicating that Bourdieusian status display dynamics may increasingly be at play for this pattern of taste. Finally, while categorical tolerance is not polarized along liberal/conservative lines, \textit{symbolic exclusion} patterns net of categorical tolerance are. After adjusting for other socio-demographic predictors, conservatives are more likely to express a greater number of cultural dislikes than liberals. This indicates that political ideology, along with age---but not education---is now the most reliable predictor of expressing dislikes for cultural genres in the United States today.

The rest of the paper is organized as follows. The next section briefly reviews recent work that builds on L\&S's idea of categorical tolerance, as well as other work examining the growing linkages between cultural tastes and political ideology. The next sections introduce the data and outline the basic analytic plan, presenting the main results. Finally, the discussion section summarizes the key findings, highlights their main implications, and suggests avenues for future work.

@@

The general empirical implication is that cultural omnivores should be more (explicitly) tolerant, liberal, and inclusive, displaying a fundamental openness to consume and engage in a wide variety of cultural practices and experiences. However, the Petersonian thesis does not imply that omnivores indiscriminately accept all cultural forms. Instead, subsequent research goes through some pains to emphasize that cultural omnivores still engage in ``patterned tolerance" and selective exclusion of low-status cultural forms \citep{bryson1996anything-311, lindblom2019anything-530}. This means that while cultural elites embrace a wide variety of cultural genres, they often express explicit dislikes toward a few specific genres, particularly those associated with lower-status groups \citep{bryson1996anything-311, lindblom2022growing-ded}.

Why are dislikes expressed in social surveys on cultural tastes important? As \citet{bourdieu1984distinction-835} himself noted, taste is often asserted negatively through refusal (p. 56). \citet{bryson1996anything-311} linked this dynamic to the idea of ``symbolic exclusion'' or the \textit{indirect} action of drawing a symbolic (social) boundary against particular (usually lower status) groups by distancing oneself from the cultural objects that those groups are perceived to consume \citep{lizardo2016cultural-aaa}. This recognition has led scholars to stress the critical importance of studying ``dislikes" alongside ``likes" to fully comprehend the dynamics of symbolic exclusion and the construction of cultural boundaries \citep{lizardo2015musical-8c6, lindblom2022growing-ded, oncini2023cultural-323}. This general framework has proven helpful for understanding social and cultural dynamics, even beyond the case of the United States. For example, a recent study of musical taste conducted in South Korea found that tolerant omnivores still excluded genres like ``gugak" (Korean traditional music) and ``trot," (one of the oldest forms of Korean popular music) which are preferred by lower-educated and older individuals, suggesting that the more ``cosmopolitan'' omnivores still engage in patterns of symbolic exclusion directed at older persons of lower educational status \citep[321]{kim2020anything-71b}.

@@

Various studies conducted since Bryson's consistently show an empirical link between cultural openness and social tolerance across multiple contexts. For instance, \citet{cho2023are-61e} show that cultural omnivores in South Korea are more accepting of immigrants, demonstrating a greater willingness to include them as friends, neighbors, and even marriage partners. Beyond immigration, cultural omnivorousness is positively correlated with a liberal political outlook and a higher degree of social trust, particularly towards outgroup members. Their diverse cultural engagements are also associated with a cosmopolitan outlook, support for social integration and inclusiveness, and progressive values such as the rejection of traditional gender roles and acceptance of alternative sexual lifestyles \citep{chan2019understanding-1d0}. In line with this work, A study conducted in Finland by \citep{lindblom2022growing-ded} indicates that omnivores hold very liberal views on immigration, law enforcement, and same-sex relationships, while prioritizing environmental protection \citep{lindblom2022growing-ded}.

Conversely, Univorism, a concept first conceptualized by \citet{peterson1992understanding-ce3} and empirically studied by \citet{bryson1997what-933}, describes individuals characterized by narrow, limited cultural tastes, typically confining themselves to a narrower range of popular cultural practices \citep{lindblom2022growing-ded, cho2023are-61e}. This restricted taste profile is often characteristic of older age groups and those with lower educational attainment \citep{sokolova2020does-65c, lindblom2019anything-530}. Recent work shows that univores exhibit less liberal and less tolerant attitudes across various social and political domains. For instance, univores tend to be positioned on the conservative end of the liberal-conservative spectrum, are more skeptical towards fellow citizens, indicating lower social trust,  have lower voter turnout, and espouse more conservative political views that are critical of immigration and generally do not prioritize environmental protection \citep{lindblom2022growing-ded, cho2023are-61e, chan2019understanding-1d0}.

@@

\section*{Data}
The data for this study come from three rounds of survey data collection on Americans' cultural tastes. The two waves of data collection yielded two convenience samples: one collected in 2021 and the other in 2023, with an oversample of people with advanced educational degrees. Both of these samples come from the Prolific Academic (ProA) platform. Prolific has emerged as a preferred alternative to Amazon’s Mechanical Turk for researchers seeking high-quality, suitably motivated participant pools. Unlike MTurk, ProA was explicitly designed with social-scientific researchers in mind and offers a high-quality, better-compensated participant pool. Also, unlike MTurk, ProA mandates a minimum payment based on the expected time required to complete the task. Participants in our study were compensated at a rate of \$9.50 per hour, which ProA indicates as a “good” compensation rate. The main participant pool for this study on the Prolific platform was limited to participants between the ages of eighteen and eighty-five who reside in the United States and whose first language is English (or are fluent in English). The study was also restricted to participants who scored at or above 95\% on ProA’s ``approval rate'' (the proportion of times a participant’s study response has been accepted in other studies as adequate). Survey completion was verified by embedding a completion code as the last question, which participants had to input on the Prolific website to confirm they had read and completed each survey question.

In the summer of 2025, I collected a third sample to replicate the results of the first two waves using a more representative sample of American respondents. For this purpose, I used the recently launched Cint (previously Lucid) Theorem platform, which provides access to a quota-sampled set of respondents matched to the U.S. census by age, gender, ethnicity, and region. Like Prolific, Cint Theorem is designed primarily for social science research. Cint employs a "marketplace" approach to recruiting study participants by connecting researchers with various suppliers and offering a variety of incentives to respondents. In this way, Cint provides relatively affordable access to participants for researchers (about \$1.50 per completed survey) and high-quality samples \citep{crabtree2025yes-e5f}. Cint also uses various automated screens (e.g., invalid traffic, duplicate IPs) and response quality indicators (e.g., CAPTCHA scores) to weed out bots and prevent less attentive participants from accessing the survey. Representative samples of the U.S. population from Cint Theorem have been used effectively in recent social scientific work on Americans' political and social attitudes (e.g., Keskintürk et al. 2025).

@@

\section*{Results}
\subsection*{Probability of Dislikes by Genre}
Table~\ref{tab:dis-by-genres} shows the probability of any genre experiencing a judgment of dislike for both the Prolific and Lucid samples, pooling the 2021 and 2023 Prolific samples. As we can see, Metal is the genre most likely to be disliked in both samples, consistent with previous work on Americans' musical tastes based on data collected earlier \citep{bryson1996anything-311, lizardo2015musical-8c6}. Both samples also agree that the two genres least likely to be disliked are Rock and Pop music. There is some disagreement across samples on the exact ranking of other genres. Country music experiences the most significant disparity, ranking second among the most disliked genres in the Prolific sample but near the bottom in the Lucid sample. Beyond that, both datasets agree that Musicals, Rap, and Disco are at the top of the dislike list, while Classical, R\&B, and Jazz are closer to the bottom.

@@

\subsection*{Univariate Distribution of Number of Dislikes}
Table~\ref{tab:dis-dist} shows the univariate distribution of the number of dislikes for both the Prolific and Lucid samples. As we can see, both datasets closely agree on the proportion of categorical tolerants, with the Prolific data yielding about 19\% and the Lucid dataset about 20\%. When we combine both, we get an estimate of 19.6\%, suggesting that about a fifth of the American population today can be classified as categorical tolerants, a presumed increase from L\&S's estimate of about 15\% using data from 2012.\footnote{We should be cautious when making direct comparisons, given that the number of genres included is smaller (twelve versus twenty in L\&S's 2012 survey) in both the Prolific and Lucid samples. Nevertheless, we can conclude that categorical tolerance remains a pervasive phenomenon among a substantial number of Americans.}

@@

\begin{figure}[ht!]
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol.png}
    \caption{Univariate plot of number of dislikes, joint Prolific and Lucid samples (N = 2367).}
    \label{fig:main}
\end{figure}

@@

\subsection*{Multivariable Analysis Results}
In their original paper, \citet{lizardo2016end-4fb} proposed modeling categorical tolerance and symbolic exclusion jointly as a hurdle process. The basic idea is that categorical tolerants are a distinct slice of the population who are not at risk of engaging in symbolic exclusion (expressing dislikes). Thus, we can statistically model the joint predictors of both tendencies by assuming that there is a set of predictors that drive the probability of being in the categorical tolerant group (thus not being at risk of expressing dislikes), and net of this, there is a second set of predictors that drive the number of dislikes expressed for those who are at risk of doing so. The zero-inflated Poisson model is thus a natural statistical framework for modeling this dual regime data-generating process \citep{zorn1998analytic-6b1}. Since the univariate distribution of cultural dislikes is comparable across samples (see ~\ref{tab:dis-dist}), I pool the three samples to increase statistical power and report results from the pooled sample analysis, including a binary indicator for each survey sample (using the 2021 Prolific sample as the reference category) in each model. I use listwise deletion of cases with missing values for the socio-demographic covariates included in the model, which leaves us with a total analytical sample of $N = 2291$.

@@

In line with this analytic strategy, Table~\ref{tab:zinf} shows the coefficient estimates from a set of four zero-inflated Poisson regression models. The specifications in each model, for both the count and zero-inflated components, are informed by the descriptive patterns we considered in Figures~\ref{fig:cat-tol} and~\ref{fig:grd-int}. Coefficient estimates corresponding to the count part of the model---predicting the number of dislikes among those who are at risk of expressing them---are shown in the top half of the table. Coefficient estimates for the zero-inflation part of the model, predicting the odds of landing in the categorical tolerant group---and thus not being at risk of engaging in symbolic exclusion---are presented in the lower half of the table. In all models, age, education, income, and political ideology (with higher values indicating a more liberal stance) are entered as ordered numerical inputs. Race (White) and Gender (Female) are entered into the model as dummy indicators. The zero-inflated specifications also contain squared terms for education and income to model expected curvilinear effects (see Figures~\ref{fig:cat-tol-edu} and~\ref{fig:cat-tol-inc}). Finally, I include indicator variables for each of the surveys (one for the 2023 Prolific sample and another for the 2025 Lucid sample), with the Prolific study conducted in 2021 as the reference category.

Let us begin with the predictors of symbolic exclusion. Model one includes all of the main socio-demographic predictors. The results align with expectations derived from the descriptive patterns considered earlier. Age increases the likelihood of engaging in symbolic exclusion, but education and income reduce this tendency. Nevertheless, the education effect, as suggested by Figure~\ref{fig:cat-tol-edu}, is substantively small and not statistically significant ($p = 0.09$) using the traditional cutoff value of $p < 0.05$, and becomes smaller and crosses the threshold to decidedly not statistically significant in later specifications. Being politically liberal, on the other hand, has a substantively significant effect, one that is robust---in terms of statistical significance---across models. As suggested by Figure~\ref{fig:cat-tol-pol}, self-identified Liberals are less likely to engage in symbolic exclusion than conservatives. These results indicate that in the United States today, political liberalism is a more reliable predictor of whether a person will engage in symbolic exclusion than is educational attainment, a situation quite different from that which obtained in the 1990s \citep{bryson1996anything-311}. When it comes to ethnoracial identity, self-identified white respondents are more likely to express dislikes compared to respondents who identified with one of the non-white ethnoracial categories ($p < 0.01$). Finally, there is no appreciable gender-identity difference in the propensity to engage in symbolic exclusion via cultural dislikes ($p = 0.53$).

Model 1 includes age as the sole demographic predictor in the zero-inflated part of the model, providing strong evidence that categorical tolerance is primarily a cultural habit among the young and is less prevalent among older respondents ($p < 0.001$). Model 2 includes race and gender as additional predictors of categorical tolerance, verifying the expectation that white respondents are less likely to be found in this group (compared to non-white respondents) and women are less likely to be categorical tolerants (compared to respondents who identify as men). Model 3 introduces education as an additional predictor of categorical tolerance, including a linear input for the respondent's education and a square term for the same covariate. The results verify the idea, suggested by Figure~\ref{fig:cat-tol-edu}, the education effect on the odds of being a categorical tolerant is parabolic, with a negative linear component (indicating that lower education person are more likely to be in the categorical tolerant group) and a positive squared component (indicating that at some point the negative effect reverses, and very high-education individuals end up over-represented among the categorical tolerants).

That said, the linear component of education is of only marginal statistical significance in Model 3 ($p = 0.051$). This suggests that the bulk of the systematic effect of education is contained in the positive quadratic component, indicating a larger propensity of the very highly educated to be over-represented among categorical tolerants. This impression is confirmed in Model 4, which adjusts for the non-linear effect of self-reported family income on the odds of being a categorical tolerant. Once we do that, the main negative linear effect of education is attenuated ($p = 0.18$), but the exponential positive effect remains ($p < 0.05$), using statistical significance as a criterion. Both components of the income effect, on the other hand, have a statistically significant impact on the odds of being classified as a CT, with a negative linear effect and a positive quadratic effect, suggesting an under-representation of people in middle-income categories among the categorically tolerant.

@@

\subsection*{Summary of Key Results}

The findings presented in this paper offer a critical update to recent work on changing cultural taste patterns in the United States, particularly regarding the dynamic processes of inclusion and exclusion in cultural preferences centered around the idea of Categorical Tolerance (CT). The core empirical results further establish the enduring relevance and continuous difussion of CT, a cultural pattern first identified in earlier work by L\&S \citeyearpar{lizardo2016end-4fb}. Specifically, approximately twenty percent of respondents across the samples systematically refused to report disliking any of the measured cultural genres, a larger proportion than that observed in previous work. This basic descriptive result confirms that the univariate distribution of dislikes maintains the expected ``zero-inflated'' structure, diverging from a simple Poisson process.
